### 2016-10-12 thesis_paper/source.tex
Formatted some thesis paper parts

Srinivas Katkoori, Ph. D.

Sameer Varma, Ph. D.

@@

October 10, 2016

@@

\hspace{3em} In collaboration with Physics Department, the original motivation came from the need of improving analytic query executions performance over big data being generated by simulation software. The idea was to take input stream coming from simulation software (ideally live) and run network of queries very efficiently, because sometimes they might take days and more to compute. The proposed improved version of push-based system for molecular simulation data analysis, we believe, gives an opportunity for scientists to run their analysis fast having an ability to expand network of queries and customize computation functions to get the desired results. 
Our system primarily represents a system that works in several modules that were developed in C, CUDA, Python programming languages that, in combination, give us the desired result to perform efficient computations for the network of queries developed in the previous paper \cite{mainPaper} as well as to prove the concept of optimization of push-based system by means of parallelism with no data loss. We've developed the following modules:

@@

\noindent\hspace{3em}As a plugin for optimizing computation of scientific queries against large data sets, an improved Push-based System for MS improvement should be able to handle data streams as well, in order to get real time results and saving up on time, instead of pulling everything from files later on.

\section{Big Data Frameworks}

\hspace{3em} We are potentially dealing with data that could easily get up to terabytes and petabytes in sizes, and it will be fair to apply term of big data. Traditional software tools are not able to process Big Data apparently because of huge data sets, thus, special frameworks were developed to handle them. Nowadays there are bunch of frameworks that try to solve this problem. Some of these frameworks are Hadoop \cite{hadoop}, Hadoop with GPU extension \cite{hadoopgpu}, Spark \cite{spark}, Flink \cite{flink} (this is more of a big data stream processing framework) and other ones that are possibly just custom modifications and distributions of these. The mentioned tools help dealing with big data giving an ability to process them in batches on clusters using underlying HDFS (Highly Distributed File System), and normally go along with other utilities that help to apply MapReduce concept. The relationship between our system and this concept along with its implementations is in data sets sizes. Even though we are not using or introducing any non-traditional file systems, in certain scale and problems it could be required, too. In fact, the frameworks listed above could even be used for our problem on top of our system, because, as we discuss in section \ref{sc:conclusion}, it could be extended to a distributed system for analyzing independent frames in parallel.

@@

\hspace{3em} Once the data is loaded into global memory of GPU, we have two kernels for two types of queries setup for two different streams. Beside it, we allocate some memory in GPU global memory for results, too, and keep it as an object with fields to easily pull it back when it is ready, in our case all the data sizes are static before the kernels are executed, which follows nature of GPU devices.

\hspace{3em} As it has already been mentioned in previous sections, one body functions are relatively easy to implement, since they don't require iteration through data more than once with a write approach. Our implementation of one-body functions are relatively easy to understand which is what we tried to explain further.

@@

\noindent\hspace{3em} The same CPU has been used for sequential as well as for improved GPU version during experiments with benchmarks. So, for CPU in GPU improved version refer to the same Table \ref{tb:cpuhardware}. For recall, CPU in improved version is used in-between memory loading from file and uploading it to GPU, printing the results to the user, which is not big load. Consequently, we also have GPU device. Hardware summary for GPU device is described in Table \ref{tb:gpuhardware}

@@

\hspace{3em} We have run number of experiments with different number of atoms and frames. Results vary as we choose different data sets, but we can see some tendency towards growth and, fortunately, it is very positive for our improved version of push-based system.

\subsection{Experiment Results For 1000 Atoms}

\hspace{3em} Even though we assume that majority of our cases involve huge amount of atoms (starting from hundred thousands), we still ran experiments on smaller amount of atoms to see difference in speedup relatively with the data set sizes. You can observe summary in Figures \ref{fg:exp_1000_10}, \ref{fg:exp_1000_100}, \ref{fg:exp_1000_1000}, \ref{fg:speedup_1000}. There is a minor threshold, where CPU version is faster, where workload is less than 20 on Figure \ref{fg:exp_1000_10}, but it is changed with growth of workload. That presumably is the minor overhead of loading data from CPU to GPU and number of kernels versus computation work actually needs to be done. Once there is more load on computation, strongly GPU accelerates and becomes faster.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Experiment Results For 10,000 Atoms}

@@

\subsection{Experiment Results For 100,000 Atoms}

@@

\subsection{Advanced SDH Implementation}

\hspace{3em} We implemented more advanced version of SDH algorithm taking better use of shared memory just to demonstrate what improvement and speedup could be achieved if we focus on particular query implementation.

@@

\begin{figure}
 \centerline{ \includegraphics[width=0.5\columnwidth]{images/sdhvsimproved}}
 \caption{Comparison of GPU SDH with GPU SDH Improved}
 \label{fg:gpusdhsdh}
\end{figure}

@@

\subsection{Overall Performance}

\hspace{3em} It is noticable that growth of the speedup is not stable and with certain number of atoms it drops dramatically. One of possible reasons could the fact that we've used static approach of computing grid dimensions. Beside it, in the future, it can be expanded into a dynamic version that will automatically adjust the grid dimensions choosing least overhead that it has to go with.

>>>>

Srinivas Katkoori, Ph.D.

Sameer Varma, Ph.D.

@@

August 23, 2016

@@

\hspace{3em} In collaboration with Physics Department, the original motivation came from the need of improving analytic query executions performance over big data being generated by simulation software. The idea was to take input stream coming from simulation software (ideally live) and run network of queries very efficiently, because sometimes they might take days and more to compute. The proposed improved version of push-based system for molecular simulation data analysis, we believe, gives an opportunity for scientists to run their analysis fast having an ability to expand network of queries and customize computation functions to get the desired results. 
Our system primarily represents a system that works in several modules that were developed in C, CUDA, Python programming languages that, in combination, give us the desired result to perform efficient computations for the network of queries developed in the previous paper \cite{mainPaper} as well as to prove the concept of optimization of push-based system by means of parallelism with no data loss:

@@

\noindent\hspace{3em}As a plugin for optimizing computation of scientific queries against large data sets, an improved Push-based System for MS improvement should be able to handle data streams as well, in order to get real time results and saving up on time, instead of pulling everything from files later on. For now it has been developed as a separate program that runs independently accepting input in a specified format which is described in the paper, but ideally it is supposed to be embedded right into Gromacs.

\section{Big Data Frameworks}
\hspace{3em} We are potentially dealing with data that could easily get up to terabytes and petabytes in sizes, and it will be fair to apply term of big data. Traditional software tools are not able to process Big Data apparently because of huge data sets, thus, special frameworks were developed to handle them. Nowadays there are bunch of frameworks that try to solve this problem. Some of these frameworks are Hadoop \cite{hadoop}, Hadoop with GPU extension \cite{hadoopgpu}, Spark \cite{spark}, Flink \cite{flink} (this is more of a big data stream processing framework) and other ones that are possibly just custom modifications and distributions of these. The mentioned tools help dealing with big data giving an ability to process them in batches on clusters using underlying HDFS (Highly Distributed File System), and normally go along with other utilities that help to apply MapReduce concept. The relationship between our system and this concept along with its implementations is in data sets sizes. Even though we are not using or introducing any non-traditional file systems, in certain scale and problems it could be required, too. In fact, the frameworks listed above could even be used for our problem on top of our system, because, as we discuss in section \ref{sc:conclusion}, it could be extended to a distributed system for analyzing independent frames in parallel.

@@

\noindent\hspace{3em} Once the data is loaded into global memory of GPU, we have two kernels for two types of queries setup for two different streams. Beside it, we allocate some memory in GPU global memory for results, too, and keep it as an object with fields to easily pull it back when it is ready, in our case all the data sizes are static before the kernels are executed, which follows nature of GPU devices.

\noindent\hspace{3em} As it has already been mentioned in previous sections, one body functions are relatively easy to implement, since they don't require iteration through data more than once with a write approach. Our implementation of one-body functions are relatively easy to understand which is what we tried to explain further.

@@

\noindent\hspace{3em} The same CPU has been used for sequential as well as for improved GPU version during experiments with benchmarks. So, for CPU in GPU improved version refer to the same Table \ref{tb:cpuhardware}. For recall, CPU in improved version is used in-between memory loading from file and uploading it to GPU, printing the results to the user, which is not big load. 
\clearpage
Consequently, we also have GPU device with its own performance capacities. Hardware capacities summary for GPU device has been described in Table \ref{tb:gpuhardware}

@@

\noindent\hspace{3em} We have run number of experiments with different number of atoms and frames. Results vary as we choose different data sets, but we can see some tendency towards growth and, fortunately, it is very positive for our improved version of push-based system.

\noindent\hspace{3em} Even though we assume that majority of our cases involve huge amount of atoms (starting from hundred thousands), we still ran experiments on smaller amount of atoms to see difference in speedup relatively with the data set sizes. You can observe summary in Figures \ref{fg:exp_1000_10}, \ref{fg:exp_1000_100}, \ref{fg:exp_1000_1000}, \ref{fg:speedup_1000}. There is a minor threshold, where CPU version is faster, where workload is less than 20 on Figure \ref{fg:exp_1000_10}, but it is changed with growth of workload. That presumably is the minor overhead of loading data from CPU to GPU and number of kernels versus computation work actually needs to be done. Once there is more load on computation, strongly GPU accelerates and becomes faster.

@@

@@

@@

\noindent\hspace{3em} We implemented more advanced version of SDH algorithm taking better use of shared memory just to demonstrate what improvement and speedup could be achieved if we focus on particular query implementation.

@@

\begin{figure}
 \centerline{ \includegraphics[width=0.5\columnwidth]{images/sdhvsimproved}}
 \caption{Comparison of GPU SDH with GPU SDH improved}
 \label{fg:gpusdhsdh}
\end{figure}

@@

\noindent\hspace{3em} It is noticable that growth of the speedup is not stable and with certain number of atoms it drops dramatically. One of possible reasons could the fact that we've used static approach of computing grid dimensions. Beside it, in the future, it can be expanded into a dynamic version that will automatically adjust the grid dimensions choosing least overhead that it has to go with.
